---
title: "R Course"
author: "R Spitzer"
date: "10/10/2018"
output: 
  html_document: 
    code_folding: hide
    highlight: monochrome
    keep_md: yes
    theme: cosmo
    toc: yes
    toc_depth: 5
    toc_float: yes
editor_options: 
  chunk_output_type: console
---
**Set Java**
```{r}
#Sys.setenv(JAVA_HOME="C:\\Program Files\\Java\\jre1.8.0_181")
Sys.setenv(JAVA_HOME="C:/Program Files/Java/jre1.8.0_181")
#Sys.getenv("JAVA_HOME")
```

**Libraries**
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(xlsx)
```

#Data
```{r}
#data.df <- read.csv(file.path("C:\\Robert\\testrun\\data","gapminder_data.csv"))
data.df <- read_csv("data/gapminder_data.csv")
```

#Playing with the data
```{r}
str(data.df)

data.df_f <- data.df %>% 
  mutate_if(sapply(data.df, is.character), as.factor)

glimpse(data.df)
glimpse(data.df_f)
```

#DF and a List
```{r}
names <- c("Robert","Linda","Julia")
my_df <- data.frame(f_names = names,
                    my_nums = c(30,75,50),
                    my_lgl = c(TRUE,FALSE,TRUE),
                    stringsAsFactors = F)
my_df
```

#Dplyr
```{r}
data.df_f %>% filter(year == 1952) %>% 
  select(country, population = pop) %>% 
  arrange(population)

data.df_f %>% filter(year == 1952) %>% 
  select(country, population = pop) %>% 
  arrange(desc(population))

#National GDP pop*gdpPercap
data.df_f %>% mutate(N_GDP = pop*gdpPercap) -> data.df_f

data.df_f 
```

#Summaries
```{r}
data.df_f %>% filter(year==2007) %>% 
  group_by(continent) %>% 
  summarise(AvgLE=mean(lifeExp),
            SD = sd(lifeExp))

#Summarize Median gdp by continent in 2007
data.df_f %>% filter(year==2007) %>% 
  mutate(N_GDP = pop*gdpPercap) %>% 
  group_by(continent) %>% 
  summarise(MedianGDP=median(N_GDP))

data.df_f %>% filter(continent=="Oceania") %>% 
  distinct(country)
```

#Functions

##if / else
```{r}
x <- 15

if(x>=10){
  paste("x is greater than or equal to 10; it is:",x)
} else {
  print("x is less than 10")
}
```

##loops
```{r}
for(i in 1:10) {
  print(i)
}
```

##nested loop
```{r}
for(i in 1:5) {
  for(j in c("a","b","c","d","e")) {
    print(paste(i,j))
    
  }
}
```

##Roll dice function
```{r}
#Step 1 (Menu -> Code -> extract function)
##Function to roll 6 sided dice only
roll <- function() {
  die <- 1:6
  dice <- sample(die, size = 2, replace = T) # size=2, sample twice
  sum(dice)
}

#roll()

#Step 2 Function to roll two 6-sided dice of any size
roll_n <- function(die_n = 1:6) {
  dice <- sample(die_n, size = 2, replace = T)
  sum(dice)
}

roll_n(1:15)
```

##Load dice function using source
```{r}
source("data/dice-function.R")
roll_n()

#source("path/dice-function.R")
```

##Gapminder function to get GDP for specific Year and/or Country
```{r}
calcGDP <- function(dat, input_yr = NULL, ctry = NULL)
{
  #Is there a year argument?
  if(!is.null(input_yr)) {
    dat <- dat %>% 
      filter(year == input_yr)
  }
  #Is there a country argument?
  if(!is.null(ctry)) {
    dat <- dat %>% 
      filter(country == ctry)
  }
  new_dat <- dat %>% 
    mutate(gdp = pop*gdpPercap)
  return(new_dat)
}

calcGDP(data.df, input_yr=1952, ctry = "Sweden")
Ger_1952 <- calcGDP(data.df, input_yr=1952, ctry = "Germany")
View(Ger_1952)

calcGDP(data.df, input_yr = 1952) %>% View()
calcGDP(data.df, NULL, "Germany") %>% View()
```

#Plots
```{r}
ggplot(data.df_f, aes(continent,lifeExp)) +
  geom_boxplot() +
  theme_bw()

ggplot(data.df_f, aes(continent,gdpPercap)) +
  geom_boxplot() +
  theme_bw()

ggplot(data.df_f, aes(log(gdpPercap),log(N_GDP), color=continent)) +
  geom_point() +
  theme_bw()

ggplot(data.df_f, aes(year,log(N_GDP), color=continent)) +
  #geom_point() +
  stat_smooth(method = 'loess') +
  theme_bw()

ggplot(data.df_f, aes(factor(year),log(N_GDP))) +
  geom_boxplot() +
  stat_smooth(aes(group=1),method = 'loess') +
  theme_bw() +
  facet_wrap(~continent)
```

#...more Plots
```{r}
data.df_f %>% ggplot(aes(year,log(N_GDP), color=continent)) + 
  stat_smooth(method = 'loess') +
  theme_bw()
```

Plot populations European countries in 1997 as bar plot
```{r}
data.df_f %>% filter(continent=="Europe" & year==1997) %>% 
  ggplot(aes(country,pop)) + 
  geom_bar(stat="identity") + 
  geom_smooth(aes(group=1),method = "loess", se=F) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Re-plot with ordering country according to population and add life expectancy
```{r}
data.df_f %>% filter(continent=="Europe" & year==1997) %>% 
  ggplot(aes(x= fct_reorder(country,pop), y=pop, fill=lifeExp)) + 
  geom_bar(stat="identity") +
  coord_flip() +
  theme_bw()
```

```{r}
data.df_f %>% ggplot(aes(fct_reorder(continent,lifeExp),lifeExp)) +
  geom_boxplot(outlier.color = "blue") +
  geom_jitter(position=position_jitter(width = .1,
                                       height = 0),
              alpha=.25) +
  theme_bw()
```

```{r}
data.df_f %>% filter(continent!="Oceania") %>% ggplot(aes(year, lifeExp, color=country)) +
  geom_line() +
  theme_classic() +
  facet_wrap(~continent) +
  theme(legend.position="none")
```











